---
title: "LPA_trial"
author: "Bowen Liu"
date: "6/12/21"
output: html_document
---

## PSS

A more reasonable way is to include 1738 persons that completes the PSS-10 survey. Then claim that 1699 of them complete "all" the corresponding demographic questions. (Which, we avoid the imputation at this stage. (39/1738 did not enter all the responses))
```{r PSS: Perceived Stress Scale}
setwd("~/Desktop/covid19_survey")
covid = read.csv("covid19_data.csv")
sum(is.na(covid$Q1)) 
# There are 2 null responses for Q1, which indicates these 2 "participants" did 
# not click on anaything to the informed consent form at the beginning. Hence, it's reasonabale to exclude them as part of this study. 
covid = covid[-c(which(is.na(covid$Q1)==TRUE)),] 
dim(covid) #Check if the null responses have been removed. 
sum(covid$Q1 == 6) 
#6 indicates that these "participants"" answered no for the informed consent form, 
#Thus, we have 5 people did not agree with the information consent form. They
#should be excluded from the participants of this study. 
covid = covid[-c(which(covid$Q1 == 6)),] 
#dim(covid) #Now this should leave us with 83 responses in total from the start
#### Demographics continues here
covid_pss = covid[-c(which(is.na(covid$PSS_1)==TRUE|is.na(covid$PSS_2)==TRUE|is.na(covid$PSS_3)==TRUE|
is.na(covid$PSS_4)==TRUE|is.na(covid$PSS_5)==TRUE|is.na(covid$PSS_6)==TRUE|is.na(covid$PSS_7)==TRUE|is.na(covid$PSS_8)==TRUE|is.na(covid$PSS_9)==TRUE|is.na(covid$PSS_10)==TRUE)),]
 #345 responses was deleted here. 1738 responses left for CAS scale calculation
##### We shall only include complete responses for a scale. 
##### Adding a column for total score of PSS, notice in the qualtrics the responses were labelled as 1-5.
##### Instead the actual range for each response is 0-4. 
##### 1,2,3,6,9,10 are negative questions, so sum them first and -6 for the reasons stated above. 
##### For 4,5,7,8, the original range in the dataset is 1-5. Now we reverse it by using 5 minus the item. 
covid_pss_scale = covid_pss[,c(19:28)]
library(psych)
KMO(covid_pss_scale)
bartlett.test(covid_pss_scale)

attach(covid_pss)
covid_pss$PSS_total = (PSS_1+PSS_2+PSS_3+ PSS_6+PSS_9+PSS_10 -6)+(5-PSS_4)+(5-PSS_5)+(5-PSS_7)+(5-PSS_8) 
detach(covid_pss)

#### FA
covid_pss_scale = covid_pss
covid_pss_scale$PSS_1 = covid_pss_scale$PSS_1-1
covid_pss_scale$PSS_2 = covid_pss_scale$PSS_2-1
covid_pss_scale$PSS_3 = covid_pss_scale$PSS_3-1
covid_pss_scale$PSS_6 = covid_pss_scale$PSS_6-1
covid_pss_scale$PSS_9 = covid_pss_scale$PSS_9-1
covid_pss_scale$PSS_10 = covid_pss_scale$PSS_10-1
covid_pss_scale$PSS_4 = 5-covid_pss_scale$PSS_4
covid_pss_scale$PSS_5 = 5-covid_pss_scale$PSS_5
covid_pss_scale$PSS_7 = 5-covid_pss_scale$PSS_7
covid_pss_scale$PSS_8 = 5-covid_pss_scale$PSS_8
covid_pss_scale$total = covid_pss$PSS_total
#covid_pss_men = subset(covid_pss_scale,Q1.1 == 1)
#covid_pss_women = subset(covid_pss_scale,Q1.1 == 2)
covid_pss_scale = covid_pss_scale[,c(19:28)]
#covid_pss_scale_women = covid_pss_women[,c(19:28)]
#covid_pss_scale_men = covid_pss_men[,c(19:28)]
library(MASS)
library(psych)
library(REdaS)
KMO(covid_pss_scale)
#Diagnol of the anti-image correlation matrix
diag(KMO(covid_pss_scale)$ImCov)
###Only 4,5,7,8,9 are included.(Since the corresponding diagonal elements of the anti-image correlation matridx > 0.5)
covid_pss_scale = covid_pss_scale[,c(4,5,7,8,9)]
#bartlett.test(covid_pss_scale)
bart_spher(covid_pss_scale)
KMO(covid_pss_scale)

pca_pss = princomp(covid_pss_scale, scores=TRUE, cor=TRUE)
screeplot(pca_pss, type="line", main="Scree Plot") 

summary(pca_pss)
library(nFactors)
fa.parallel(covid_pss_scale)
a = factanal(covid_pss_scale,factor=2,rotation = "varimax",scores = "regression")
communality = 100*(1 - a$uniquenesses) 
fa.pss <- fa(r=covid_pss_scale, 
 nfactors = 2, SMC=TRUE, covar=TRUE,
 fm="pa", # type of factor analysis we want to use (“pa” is principal axis factoring)
 max.iter=50, # (50 is the default, but we have changed it to 100
 rotate="varimax") # none rotation
print(fa.pss)
fa.diagram(fa.pss)
factor1 = fa.pss$scores[,1]
factor2 = fa.pss$scores[,2]
IQR(factor1)
length(which(factor1<quantile(factor1,0.75)+1.5*IQR(factor1) & factor1>quantile(factor1,0.25)-1.5*IQR(factor1)))
include_factor = which(factor1<quantile(factor1,0.75)+1.5*IQR(factor1) & factor1>quantile(factor1,0.25)-1.5*IQR(factor1))

#length(which(factor2<quantile(factor2,0.75)+1.5*IQR(factor2) & factor1>quantile(factor2,0.25)-1.5*IQR(factor2)))

sex = covid_pss$Q1.1
race = covid_pss$Q4
fin = covid_pss$Q24
class = covid_pss$Q6
age = covid_pss$Q5
employ = covid_pss$Q11
marital = covid_pss$Q3
total = covid_pss$PSS_total
# We want to set the base levels correspondingly
# sex: transgender as 0
sex[sex==3] = 0
#race: other as 0
race[race==9] = 0
#fin: no change as 0
fin[fin==3] = 0
#class: other as 0
class[class==9] = 0
#age: 75 or older as 0
## age 3,4,5,6,7 combine
## There is a signficant change since the base level changed in this analysis
age[age==3] = 0
age[age==4] = 0
age[age==5] = 0
age[age==6] = 0
age[age==7] = 0
# This is expected to be 57 responses from the "45 or over group". 
# The univariate analysis needs to be done again
# The base level listed in the manuscript is incorrect, but this can be explained since we 'redo the analysis'
# check the stepwise regression part.

#employ: other as 0 
employ[employ==7] = 0
#marital: other as 0
marital[marital==7] = 0
#child = covid_pss$Q15
factor_score_anal = data.frame(class = as.factor(class),
                               sex = as.factor(sex),
                               race = as.factor(race),
                               fin = as.factor(fin),
                               age = as.factor(age),
                               employ = as.factor(employ),
                               marital = as.factor(marital),
                               #child = as.factor(child),
                               factor1 = factor1,
                               factor2 = factor2,
                               total = total
                               
                          )

factor_score_anal = na.omit(factor_score_anal) #1699 left
#univariate
tapply(factor_score_anal$total,factor_score_anal$class,mean)
tapply(factor_score_anal$total,factor_score_anal$sex,mean)
tapply(factor_score_anal$total,factor_score_anal$race,mean)
tapply(factor_score_anal$total,factor_score_anal$fin,mean)
tapply(factor_score_anal$total,factor_score_anal$age,mean)#notice '0' stands for '45 and over'
tapply(factor_score_anal$total,factor_score_anal$employ,mean)
tapply(factor_score_anal$total,factor_score_anal$marital,mean)
# ANOVA: H_0: The mean score is the same for all the group
summary(aov(total~class,data = factor_score_anal)) 
summary(aov(total~sex,data = factor_score_anal))
summary(aov(total~race,data = factor_score_anal)) 
summary(aov(total~fin,data = factor_score_anal)) 
summary(aov(total~age,data = factor_score_anal))
summary(aov(total~employ,data = factor_score_anal))
summary(aov(total~marital,data = factor_score_anal))
# Kruskal-Wallis: H_0: The rank of the groups are the same (Given the shape similarity conditions, this is equivalent to test all the groups have the same median)
kruskal.test(total~class,data = factor_score_anal) 
kruskal.test(total~sex,data = factor_score_anal)
kruskal.test(total~race,data = factor_score_anal) 
kruskal.test(total~fin,data = factor_score_anal) 
kruskal.test(total~age,data = factor_score_anal)
kruskal.test(total~employ,data = factor_score_anal)
kruskal.test(total~marital,data = factor_score_anal)

######
summary(factor_score_anal)
full.model1 = lm(factor1~class+sex+race+fin+age+employ+marital,data = factor_score_anal)
full.model2 = lm(factor2~class+sex+race+fin+age+employ+marital,data = factor_score_anal)
#plot(resid(full.model1))
step.model1 <- stepAIC(full.model1, direction = "both", 
                      trace = FALSE)
step.model2 <- stepAIC(full.model2, direction = "both", 
                      trace = FALSE)

summary(step.model1)

summary(step.model2)
mean(factor_score_anal$total)
sd(factor_score_anal$total)
stress_class = c(sum(factor_score_anal$total<=13),sum(factor_score_anal$total>13 & factor_score_anal$total<27),sum(factor_score_anal$total>=27))

### PSS score distribution
aggregate(factor_score_anal$total, list(factor_score_anal$sex), FUN=mean) 
aggregate(factor_score_anal$total, list(factor_score_anal$sex), FUN=sd) 

aggregate(factor_score_anal$total, list(factor_score_anal$class), FUN=mean) 
aggregate(factor_score_anal$total, list(factor_score_anal$class), FUN=sd) 


aggregate(factor_score_anal$total, list(factor_score_anal$race), FUN=mean) 
aggregate(factor_score_anal$total, list(factor_score_anal$race), FUN=sd) 


aggregate(factor_score_anal$total, list(factor_score_anal$fin), FUN=mean) 
aggregate(factor_score_anal$total, list(factor_score_anal$fin), FUN=sd) 


aggregate(factor_score_anal$total, list(factor_score_anal$age), FUN=mean) 
aggregate(factor_score_anal$total, list(factor_score_anal$age), FUN=sd)


aggregate(factor_score_anal$total, list(factor_score_anal$employ), FUN=mean) 
aggregate(factor_score_anal$total, list(factor_score_anal$employ), FUN=sd)

aggregate(factor_score_anal$total, list(factor_score_anal$marital), FUN=mean) 
aggregate(factor_score_anal$total, list(factor_score_anal$marital), FUN=sd) 

### The following shall be igonored since the subset selection is not required. 

#sub.model1 <- regsubsets(factor1~class+sex+race+fin+age+employ+marital,data = factor_score_anal)
#sub.model2 <- regsubsets(factor2~class+sex+race+fin+age+employ+marital,data = factor_score_anal)
#library(gglasso)
#factor_pred = data.matrix(factor_score_anal[,1:7]) #Save the predictors
#result1 = gglasso(factor_pred,factor_score_anal$factor1)
#plot(result1,log.l=FALSE)
```
#In order to use the correct variable selection approach, we need to establish necessary dummy variables and use the stepwise selection method. 
```{r}
dim(factor_score_anal)
#Dem1: Sex
#Base level: transgender
male = ifelse(factor_score_anal$sex == 1,1,0)
female = ifelse(factor_score_anal$sex == 2,1,0)
#trans = ifelse(factor_score_anal$sex == 3,1,0)
which(factor_score_anal$sex == 1) == which(male == 1) ##Validate the ifelse statement
factor_score_anal = cbind(factor_score_anal,male,female)
### Check the head again
head(factor_score_anal)

#Dem2: race
### 1. Native Alaskan/ American Indian
### 2. Asian or Asian American
### 3. Black or African American
### 4. Hispanic/Latino/a/x
### 5. MENA/Middle Eastern Origin
### 6. Pacific Islander/Hawaiian
### 7. White/ Caucasian
### 8. Biracial/Multiracial
### 9. Other
#Based level: Other
native = ifelse(factor_score_anal$race == 1,1,0)
asian= ifelse(factor_score_anal$race == 2,1,0)
black = ifelse(factor_score_anal$race == 3,1,0)
hispanic = ifelse(factor_score_anal$race == 4,1,0)
mena = ifelse(factor_score_anal$race == 5,1,0)
pac = ifelse(factor_score_anal$race == 6,1,0)
white = ifelse(factor_score_anal$race == 7,1,0)
bimulti = ifelse(factor_score_anal$race == 8,1,0)
factor_score_anal = cbind(factor_score_anal,native,asian,black,hispanic,mena,pac,white,bimulti)

#Dem3: class standing
###### Classification (Q6 coding:6: Master; 7: PhD; 8: Not seeking for degree; 9: Other)
###### 1-5: different levels of undergrad
###### Base: other
fre = ifelse(factor_score_anal$class == 1,1,0)
sop = ifelse(factor_score_anal$class == 2,1,0)
jun = ifelse(factor_score_anal$class == 3,1,0)
sen = ifelse(factor_score_anal$class == 4,1,0)
over5 = ifelse(factor_score_anal$class == 5,1,0)
master = ifelse(factor_score_anal$class == 6,1,0)
phd = ifelse(factor_score_anal$class == 7,1,0)
non_deg = ifelse(factor_score_anal$class == 8,1,0)
factor_score_anal = cbind(factor_score_anal,fre,sop,jun,sen,over5,master,phd,non_deg)

#Dem4: Marital
#Married:1
#Widowed:2
#Divorced:3
#Separated:4
#Partnered:5
#Single:6
#Other:7 Base level
mar = ifelse(factor_score_anal$marital == 1,1,0)
wid = ifelse(factor_score_anal$marital == 2,1,0)
div = ifelse(factor_score_anal$marital == 3,1,0)
sep = ifelse(factor_score_anal$marital == 4,1,0)
par = ifelse(factor_score_anal$marital == 5,1,0)
sin = ifelse(factor_score_anal$marital == 6,1,0)
factor_score_anal = cbind(factor_score_anal,mar,wid,div,sep,par,sin)

#Dem5: Age
#18-24: 1
#25-34: 2
#35-44: 3
#45-54: 4
#55-64: 5
#65-74: 6
#75 or older: 7: base level
age1 = ifelse(factor_score_anal$age == 1,1,0)
age2 = ifelse(factor_score_anal$age == 2,1,0)
age3 = ifelse(factor_score_anal$age == 3,1,0)
age4 = ifelse(factor_score_anal$age == 4,1,0)
age5 = ifelse(factor_score_anal$age == 5,1,0)
age6 = ifelse(factor_score_anal$age == 6,1,0)
factor_score_anal = cbind(factor_score_anal,age1,age2,age3,age4,age5,age6)

#Dem6: employment
#Full-time:1
#Part-time:2
#Unemployed:3
#Laid off due to COVID-19:4
#Retired:5
#Not working due to disability:6
#Other: 7 base level
emp1 = ifelse(factor_score_anal$employ == 1,1,0)
emp2 = ifelse(factor_score_anal$employ == 2,1,0)
emp3 = ifelse(factor_score_anal$employ == 3,1,0)
emp4 = ifelse(factor_score_anal$employ == 4,1,0)
emp5 = ifelse(factor_score_anal$employ == 5,1,0)
emp6 = ifelse(factor_score_anal$employ == 6,1,0)
factor_score_anal = cbind(factor_score_anal,emp1,emp2,emp3,emp4,emp5,emp6)

#Dem7: financial status
#A lot more stressful:1
#Somewhat more stressful:2
#No change: 3 -> base level
#Somewhat less stressful:4
#A lot less stressful:5
fin1 = ifelse(factor_score_anal$fin == 1,1,0)
fin2 = ifelse(factor_score_anal$fin == 2,1,0)
fin3 = ifelse(factor_score_anal$fin == 4,1,0) #Notice here directly jump to less stressful
fin4 = ifelse(factor_score_anal$fin == 5,1,0) 
factor_score_anal = cbind(factor_score_anal,fin1,fin2,fin3,fin4)
head(factor_score_anal[,1:9]) 
#Create the new data frame with only dummies
Factor1_score_anal = factor_score_anal[,c(8,10:49)] #Now the old subjects contain 49 columns and use this 
#to analyze factor 1
Factor2_score_anal = factor_score_anal[,9:49] #Now the old subjects contain 49 columns and use this 
#to analyze factor 2
full.model1 = lm(factor1~.,data = Factor1_score_anal)
full.model2 = lm(factor2~.,data = Factor2_score_anal)
#Selection based on Akaike Information Criterion
step.model1 <- stepAIC(full.model1, direction = "both", 
                      trace = FALSE)
step.model2 <- stepAIC(full.model2, direction = "both", 
                      trace = FALSE)
summary(step.model1)
summary(step.model2)

factor_male = subset(factor_score_anal,sex==1)
factor_female = subset(factor_score_anal,sex==2)
mean(factor_male$factor1)
mean(factor_female$factor1)
mean(factor_male$factor2)
mean(factor_female$factor2)
###gg lasso exploration
library(gglasso)
factor1_pred = data.matrix(Factor1_score_anal[,2:41]) #Save the predictors
result1 = gglasso(factor1_pred,Factor1_score_anal$factor1)
plot(result1)
factor2_pred = data.matrix(Factor2_score_anal[,2:41]) #Save the predictors
result2 = gglasso(factor2_pred,Factor2_score_anal$factor2)
plot(result2)

```

```{r}
library(ggplot2)
se = function(x)sqrt(var(x)/length(x))
### Other: 0
### Male: 1
### Female: 2
sex_factor1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$sex), FUN=mean)
sex_factor2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$sex), FUN=mean)
sex_se1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$sex), FUN=se)
sex_se2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$sex), FUN=se)
### Create a dataset
gender = c(rep("Other",2), rep("Male" , 2), rep("Female",2))
condition = rep(c("Irritability" , "Anxiety") , 3)
value = c(sex_factor1[1,2], sex_factor2[1,2],sex_factor1[2,2], sex_factor2[2,2],sex_factor1[3,2], sex_factor2[3,2])
sd =c(sex_se1[1,2], sex_se2[1,2],sex_se1[2,2], sex_se2[2,2],sex_se1[3,2], sex_se2[3,2])
data = data.frame(gender,condition,value)
### Grouped 
genderf = ggplot(data, aes(fill=condition, y=value, x=gender))+geom_bar(position="dodge", stat="identity")+ylab("Factor Score")+xlab("Gender")+labs(fill = "Factor")+geom_errorbar(aes(ymin = value-sd,ymax = value+sd),width=0.2,position = position_dodge(.9))+theme(text = element_text(size=20))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggsave("gender.tiff", plot = genderf, width=12, height=8, dpi=300)

#Dem3: class standing
#### Classification (Q6 coding:6: Master; 7: PhD; 8: Not seeking for degree; 9: Other)
#### 1-5: different levels of undergrad
#### Base: other
class_factor1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$class), FUN=mean)
class_factor2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$class), FUN=mean)
class_sd1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$class), FUN=se)
class_sd2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$class), FUN=se)
### Create a dataset
class = c(rep("Other",2),rep("Freshman",2), rep("Sophomore" , 2), rep("Junior",2), rep("Senior",2), rep("5th year or more undergrad",2), rep("Master's",2),rep("PhD",2), rep("non-degree seeking",2))
condition = rep(c("Irritability" , "Anxiety") , 9)
value = c(class_factor1[1,2], class_factor2[1,2],class_factor1[2,2], class_factor2[2,2],class_factor1[3,2], class_factor2[3,2], class_factor1[4,2], class_factor2[4,2],class_factor1[5,2], class_factor2[5,2],class_factor1[6,2], class_factor2[6,2], class_factor1[7,2], class_factor2[7,2],class_factor1[8,2], class_factor2[8,2],class_factor1[9,2], class_factor2[9,2])

sd =  c(class_sd1[1,2], class_sd2[1,2],class_sd1[2,2], class_sd2[2,2],class_sd1[3,2], class_sd2[3,2], class_sd1[4,2], class_sd2[4,2],class_sd1[5,2], class_sd2[5,2],class_sd1[6,2], class_sd2[6,2], class_factor1[7,2], class_sd2[7,2],class_sd1[8,2], class_sd2[8,2],class_sd1[9,2], class_sd2[9,2])

data = data.frame(class,condition,value)
### Grouped 
classf = ggplot(data, aes(fill= class, y=value, x=condition))+geom_bar(position="dodge", stat="identity")+ylab("Factor Score")+xlab("Factor")+labs(fill = "Class Standing")+geom_errorbar(aes(ymin = value-sd,ymax = value+sd),width=0.2,position = position_dodge(.9))+theme(text = element_text(size=20))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggsave("class.tiff", plot = classf, width=12, height=8, dpi=300)


#Dem2: race
### 1. Native Alaskan/ American Indian
### 2. Asian or Asian American
### 3. Black or African American
### 4. Hispanic/Latino/a/x
### 5. MENA/Middle Eastern Origin
### 6. Pacific Islander/Hawaiian
### 7. White/ Caucasian
### 8. Biracial/Multiracial
### 0. Other
#Based level: Other

race_factor1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$race), FUN=mean)
race_factor2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$race), FUN=mean) 
race_sd1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$race), FUN=se)
race_sd2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$race), FUN=se)
Ethnicity = c(rep("Other",2),rep("Asian/Asian American",2),rep("Native American/American Indian",2), rep("Black/African American" , 2), rep("Hispanic",2), rep("Middle Eastern Origin",2), rep("Pacific Islander/Hawaiian",2), rep("White/Caucasian",2), rep("Biracial/Multiracial",2))
Factor = rep(c("Irritability" , "Anxiety") , 9)
value = c(race_factor1[1,2], race_factor2[1,2],race_factor1[2,2], race_factor2[2,2],race_factor1[3,2], race_factor2[3,2], race_factor1[4,2], race_factor2[4,2],race_factor1[5,2], race_factor2[5,2],race_factor1[6,2], race_factor2[6,2], race_factor1[7,2], race_factor2[7,2],race_factor1[8,2], race_factor2[8,2],race_factor1[9,2], race_factor2[9,2])

sd =  c(race_sd1[1,2], race_sd2[1,2],race_sd1[2,2], race_sd2[2,2],race_sd1[3,2], race_sd2[3,2], race_sd1[4,2], race_sd2[4,2],race_sd1[5,2], race_sd2[5,2],race_sd1[6,2], race_sd2[6,2], race_factor1[7,2], race_sd2[7,2],race_sd1[8,2], race_sd2[8,2],race_sd1[9,2], race_sd2[9,2])
data = data.frame(Ethnicity,condition,value)
### Grouped 
ethnicityf = ggplot(data, aes(fill=Ethnicity, y=value, x=Factor))+geom_bar(position="dodge", stat="identity")+ylab("Factor Score")+xlab("Factor")+labs(fill = "Ethnicity")+geom_errorbar(aes(ymin = value-sd,ymax = value+sd),width=0.2,position = position_dodge(.9))+theme(text = element_text(size=20))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggsave("ethnicityf.tiff", plot = ethnicityf, width=12, height=8, dpi=300)

#Dem7: financial status
###A lot more stressful:1
###Somewhat more stressful:2
###No change: 3 base level
###Somewhat less stressful:4
###A lot less stressful:5

fin_factor1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$fin), FUN=mean) 
fin_factor2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$fin), FUN=mean) 
fin_sd1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$fin), FUN=se)
fin_sd2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$fin), FUN=se)
### Create a dataset
financial_status = c(rep("no \n change",2), rep("a lot more \n stressful",2), rep("somewhat more \n stressful",2), rep("less \n stressful",2), rep("a lot less \n stressful",2))
condition = rep(c("Irritability" , "Anxiety") , 5)
value = c(fin_factor1[1,2], fin_factor2[1,2],fin_factor1[2,2], fin_factor2[2,2],fin_factor1[3,2], fin_factor2[3,2], fin_factor1[4,2], fin_factor2[4,2], fin_factor1[5,2], fin_factor2[5,2])
sd =  c(fin_sd1[1,2], fin_sd2[1,2],fin_sd1[2,2], fin_sd2[2,2],fin_sd1[3,2], fin_sd2[3,2], fin_sd1[4,2], fin_sd2[4,2],fin_sd1[5,2], fin_sd2[5,2])
data = data.frame(financial_status,condition,value)
### Grouped 
finf = ggplot(data, aes(fill=condition, y=value, x=financial_status))+geom_bar(position="dodge", stat="identity")+ylab("Factor Score")+xlab("Financial Status")+labs(fill = "Factor")+geom_errorbar(aes(ymin = value-sd,ymax = value+sd),width=0.2,position = position_dodge(.9))+theme(text = element_text(size=20))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggsave("finf.tiff", plot = finf, width=12, height=8, dpi=300)

#Dem5: Age
###18-24: 1
###25-34: 2
###35-44: 3
###45-54: 4
###55-64: 5
###65-74: 6
###75 or older: 7: base level

age_factor1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$age), FUN=mean) 
age_factor2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$age), FUN=mean)
age_sd1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$age), FUN=se)
age_sd2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$age), FUN=se)
### Create a dataset
age_group = c(rep("18-25",2), rep("25-34",2), rep("35-44",2), rep("45-54",2), rep("55-64",2),rep("65-74",2))
condition = rep(c("Irritability" , "Anxiety") , 6)
value = c(age_factor1[1,2], age_factor2[1,2],age_factor1[2,2], age_factor2[2,2],age_factor1[3,2], age_factor2[3,2], age_factor1[4,2], age_factor2[4,2], age_factor1[5,2], age_factor2[5,2],age_factor1[6,2], age_factor2[6,2])
sd =  c(age_sd1[1,2], age_sd2[1,2],age_sd1[2,2], age_sd2[2,2],age_sd1[3,2], age_sd2[3,2], age_sd1[4,2], age_sd2[4,2],age_sd1[5,2], age_sd2[5,2],age_sd1[6,2], age_sd2[6,2])
data = data.frame(age_group,condition,value)
### Grouped 
agef = ggplot(data, aes(fill=age_group, y=value, x=condition))+geom_bar(position="dodge", stat="identity")+ylab("Factor Score")+xlab("Factor")+labs(fill = "Age Group")+geom_errorbar(aes(ymin = value-sd,ymax = value+sd),width=0.2,position = position_dodge(.9))+theme(text = element_text(size=20))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggsave("agef.tiff", plot = agef, width=12, height=8, dpi=300)

###Dem6: employment
###Full-time:1
###Part-time:2
###Unemployed:3
###Laid off due to COVID-19:4
###Retired:5
###Not working due to disability:6
###Other: 7 base level

emp_factor1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$employ), FUN=mean) 
emp_factor2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$employ), FUN=mean)
emp_sd1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$employ), FUN=se)
emp_sd2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$employ), FUN=se)
### Create a dataset
emp_group = c(rep("Other",2),rep("Full-time",2), rep("Part-time",2), rep("Unemployed",2), rep("Laid off due to COVID-19",2), rep("Retired",2),rep("Not working due to disability",2))
condition = rep(c("Irritability" , "Anxiety") , 7)
value = c(emp_factor1[1,2], emp_factor2[1,2],emp_factor1[2,2], emp_factor2[2,2],emp_factor1[3,2], emp_factor2[3,2], emp_factor1[4,2], emp_factor2[4,2], emp_factor1[5,2], emp_factor2[5,2],emp_factor1[6,2], emp_factor2[6,2], emp_factor1[7,2], emp_factor2[7,2])
sd =  c(emp_sd1[1,2], emp_sd2[1,2],emp_sd1[2,2], emp_sd2[2,2],emp_sd1[3,2], emp_sd2[3,2], emp_sd1[4,2], emp_sd2[4,2],emp_sd1[5,2], emp_sd2[5,2],emp_sd1[6,2], emp_sd2[6,2],emp_sd1[7,2], emp_sd2[7,2])
data = data.frame(emp_group,condition,value)
### Grouped 
empf = ggplot(data, aes(fill=emp_group, y=value, x=condition))+geom_bar(position="dodge", stat="identity")+ylab("Factor Score")+xlab("Factor")+labs(fill = "Employment Status")+geom_errorbar(aes(ymin = value-sd,ymax = value+sd),width=0.2,position = position_dodge(.9))+theme(text = element_text(size=20))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggsave("empf.tiff", plot = empf, width=12, height=8, dpi=300)

### Dem4: Marital
### Married:1
### Widowed:2
### Divorced:3
### Separated:4
### Partnered:5
### Single:6
### Other:7 Base level

mar_factor1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$marital), FUN=mean)
mar_factor2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$marital), FUN=mean) 
mar_sd1 = aggregate(factor_score_anal$factor1, list(factor_score_anal$marital), FUN=se)
mar_sd2 = aggregate(factor_score_anal$factor2, list(factor_score_anal$marital), FUN=se)
### Create a dataset
mar_group = c(rep("Other",2), rep("Married",2), rep("Widowed",2), rep("Divorced",2), rep("Separated",2),rep("Partnered",2), rep("Single",2))
condition = rep(c("Irritability" , "Anxiety") , 7)
value = c(mar_factor1[1,2], mar_factor2[1,2],mar_factor1[2,2], mar_factor2[2,2],mar_factor1[3,2], mar_factor2[3,2], mar_factor1[4,2], mar_factor2[4,2], mar_factor1[5,2], mar_factor2[5,2],mar_factor1[6,2], mar_factor2[6,2], mar_factor1[7,2], mar_factor2[7,2])
sd =  c(mar_sd1[1,2], mar_sd2[1,2],mar_sd1[2,2], mar_sd2[2,2],mar_sd1[3,2], mar_sd2[3,2], mar_sd1[4,2], mar_sd2[4,2],mar_sd1[5,2], mar_sd2[5,2],mar_sd1[6,2], mar_sd2[6,2],mar_sd1[7,2], mar_sd2[7,2])
data = data.frame(emp_group,condition,value)
### Grouped 
marf = ggplot(data, aes(fill=mar_group, y=value, x=condition))+geom_bar(position="dodge", stat="identity")+ylab("Factor Score")+xlab("Factor")+labs(fill = "Marital Status")+geom_errorbar(aes(ymin = value-sd,ymax = value+sd),width=0.2,position = position_dodge(.9))+theme(text = element_text(size=20))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
ggsave("marf.tiff", plot = marf, width=12, height=8, dpi=300)
```

